### 2015-04-07 contract-language.tex
Added terry classification.

\begin{figure}
\vspace{-7pt}
\begin{smathpar}
\begin{array}{rclcl}
\multicolumn{5}{c}{
  {x,y,\cureff} \in \mathtt{EffVar} \qquad \qquad
  {\sf Op} \in \mathtt{OperName}
}\\
\cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
        \ALT \forall x.\cv \ALT \pi \\
\tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
        \ALT \tau \vee \tau \\
\pi			&	\in & \mathtt{Prop} & \coloneqq & \true \ALT R(x,y)
        \ALT \pi \vee \pi \\
			  & 		&	 &  \ALT & \pi \wedge \pi \ALT \pi \Rightarrow \pi \\
R				& \in & \mathtt{Relation}	& \coloneqq & \visZ \ALT \soZ
        \ALT \sameobjZ \ALT R^+ \\
				&			&	 &  \ALT & R \cup R \ALT R \cap R \\
\end{array}
\end{smathpar}
\caption{Contract language.}
\label{fig:contract-lang}
\end{figure}

@@

\begin{figure}
\vspace{-7pt}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

\begin{itemize}[noitemsep]

@@

  \vspace{-1em}
  \begin{smathpar}
  \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

@@

  \vspace{-1em}
  \begin{smathpar}
  \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

@@

  \vspace{-1em}
  \begin{smathpar}
  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
  \end{smathpar}
\end{itemize}
\vspace{-1em}

@@

\newcommand{\DDe}[1]{#1}
\begin{figure}
\vspace{-10pt}
\begin{smathpar}
\begin{array}{cc}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  &

@@

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\vspace{-3mm}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{itemize}[noitemsep]
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{smathpar}

@@

\vspace{-1em}
\begin{smathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{smathpar}

>>>>

\begin{figure}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{rclcl}
\multicolumn{5}{c}{
  {x,y,\cureff} \in \mathtt{EffVar} \qquad \qquad
  {\sf Op} \in \mathtt{OperName}
}\\
\cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
        \ALT \forall x.\cv \ALT \pi \\
\tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
        \ALT \tau \vee \tau \\
\pi			&	\in & \mathtt{Prop} & \coloneqq & \true \ALT R(x,y)
        \ALT \pi \vee \pi \\
			  & 		&	 &  \ALT & \pi \wedge \pi \ALT \pi \Rightarrow \pi \\
R				& \in & \mathtt{Relation}	& \coloneqq & \visZ \ALT \soZ
        \ALT \sameobjZ \ALT R^+ \\
				&			&	 &  \ALT & R \cup R \ALT R \cap R \\
\end{array}
\end{smathpar}
\caption{Contract language.}
\label{fig:contract-lang}
\end{figure}

@@

\begin{figure}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

\begin{itemize}

@@

  \vspace{-1mm}
  \begin{cmathpar}
  \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{cmathpar}

@@

  \vspace{-1mm}
  \begin{cmathpar}
  \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{cmathpar}

@@

  \vspace{-1mm}
  \begin{cmathpar}
  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
  \end{cmathpar}
\end{itemize}

@@

\newcommand{\DDe}[1]{#1}
\begin{figure}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{cc}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  &

@@

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{itemize}
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

\subsection{Generality of Contracts}

It is important to note that our contract language provides a generic way to
capture application-level consistency properties and is not tied to a
particular store semantics. In particular, the same application-level contracts
can easily be mapped to a different store with a varied consistency lattice. To
illustrate this, let us consider the consistency lattice proposed by Terry et
al.~\cite{Session} based on session guarantees. Terry et al. propose the
following four incomparable session guarantees, whose semantics is captured in
the contracts below:

\vspace{-2mm}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{rcl}
\R{Read Your Writes (RYW)} & \coloneqq & \forall a,b. ~\soo{a}{b} \Rightarrow \vis{a}{b} \\
\R{Monotonic Reads (MR)} & \coloneqq & \forall a,b,c. ~\vis{a}{b} \wedge \soo{b}{c} \\
 & & \quad \Rightarrow \vis{a}{c} \\
\R{Monotonic Writes (MW)} & \coloneqq & \forall a,b,c. ~\soo{a}{b} \wedge \vis{b}{c} \\
 & & \quad \Rightarrow \vis{a}{c} \\
\R{Writes Follow Reads (WFR)} & \coloneqq & \forall a,b,c,d. ~\vis{a}{b} \wedge \vis{c}{d} \\
 & & \wedge (\sooZ ~\cup =)(b,c) \Rightarrow \vis{a}{d}
\end{array}
\end{smathpar}

Here, the consistency level for an operation in this scheme is any combination
of the above guarantees. It is easy to see that there are 24 (4!) combinations,
which form a partially ordered consistency lattice with a height of 4.
Classifying a contract under this scheme amounts to determining the weakest
combination under which the contract will hold, and entails 24 comparisons at
worst. Under this scheme, \cf{deposit} operations does not need any guarantees,
\cf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
\Rightarrow \cv_{gb}$), and \cf{withdraw} cannot be satisfied ($\Delta \vdash
\R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow \cv_{w}$).

@@

\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{smathpar}

@@

\begin{smathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{smathpar}
